---
title: 'Worksheet 5: Advanced Data Manipulation'
author: "Chase McNeill"
date: "March 4th"
---

_This is the fifth in a series of worksheets for History 8500 at Clemson University. The goal of these worksheets is simple: practice, practice, practice. The worksheet introduces concepts and techniques and includes prompts for you to practice in this interactive document. When you are finished, you should change the author name (above), commit your changes, push to GitHub, and submit the link to your worksheet repository on Canvas. Don't forget to commit your changes as you go - you should make multiple commits as you work through the worksheet, not just one at the end. Also, remember to update your learning log in the `logs/` folder of your repository as you work through the assignment._

Last week you learned the basic dplyr verbs for data manipulation. This week we'll build on those skills with more advanced techniques: grouping data in complex ways, working with categorical data using factors, joining datasets together, and reshaping data between wide and long formats.

```{r}
library(DigitalMethodsData)
library(tidyverse)
data("gayguides")
data("BostonWomenVoters")
data("statepopulations")
```
## Group_by() and Summarize()

Arrange is useful for finding the highest and lowest values, but it returns those values for the entire dataset. `group_by()`, in contrast, takes an existing tbl and converts it into a grouped tbl where operations are performed "by group". Lets look at what that means in practice: 
```{r}
mydata <- gayguides %>% 
            select(title, Year) %>%
            group_by(Year)
```
It doesn't appear that this did much. But if you hover over this new variable in your environment pane, you'll see that its now listed as a "grouped data frame." Compare that to `gayguides` which is listed as just a data frame. This means that now we can run calculations on this data and it'll perform them "by group". Or, in other words, it'll perform operations on each year within the dataset. That's where `summarize()` comes in. `summarize()` creates a new data frame with one (or more) rows for each combination of grouping variables. In this case our grouping is by year, so the resulting data frame will group records by each year in the `gayguides` dataset.

```{r}
gayguides %>% 
    select(title, Year) %>%
    group_by(Year) %>%
    summarize(count = n())
```
What happened here? In this example, we asked group_by to create groups based on year and then in summarize we created a column called count. We passed it the n() function which gives the current group size. What results, is a dataset that lists each year and how many locations that state has. 

You can group by more than one variable at a time. This creates groups for every combination of those variables.
```{r}
# Group by both state AND year
gayguides %>%
  group_by(state, Year) %>%
  summarize(count = n()) %>%
  head(10)
```

This gives us the count of locations for each state in each year.

You can also create multiple summary columns in a single `summarize()` call:
```{r}
BostonWomenVoters %>%
  group_by(Occupation) %>%
  summarize(
    count = n(),
    avg_age = mean(Age, na.rm = TRUE),
    min_age = min(Age, na.rm = TRUE),
    max_age = max(Age, na.rm = TRUE)
  ) %>%
  head()
```

**Note:** The `na.rm = TRUE` argument tells R to remove NA values before calculating. This is important when working with historical data that may have missing values.

(@) You try, use group_by and summarize to find the total number of locations in each state, each year.
```{r}
state_year_gayguides <- gayguides |>
  group_by(state, Year) |>
  summarize(count = n())
gayguides |>
  group_by(state, Year) |>
  summarize(count = n())
```

(@) Summarize can do more than just count rows. Can you use `summarize()` to find the average age for each occupation in the Boston Women Voters data?
```{r}
occupation_age <- BostonWomenVoters |>
  group_by(Occupation) |>
  summarize(
    count = n(),
    avg_age = mean(Age, na.rm = TRUE),
  ) 
```

(@) In the `gayguides` data, on average how many locations did each city in South Carolina have between 1970 and 1975?
```{r}
seventy_gayguides <- gayguides |> filter(1970 <= Year & Year <= 1975) |> group_by(state, Year) |> summarize(count = n())
```

(@) Can you group the gayguides data by state AND type, then count how many of each type of location exists in each state? Which state has the most bars?
```{r}
state_type <- gayguides |> group_by(state, type) |> summarize(count=n()) |> arrange(desc(count))
head(state_type, 1)
##California 
```

(@) Filter the dataset for only the values in the southernstates list (created in the block below). Then tell me, how many locations were in all the southern states in 1975?
```{r}
southernstates <- c("AL", "AR", "FL", "GA", "KY", "LA", "MD", "MS", "NC", "SC", "TN", "TX", "VI", "WV")
southern_gay_guides <- gayguides |> filter(state == southernstates) |> filter(Year == 1975) 
nrow(southern_gay_guides)
## 44
```


## Re-Shaping Data: Joins and Pivots

### Joins()
At some point, you might have a situation where you want to join two tables together. For example, in the `almshouse_admissions` dataset there is a column called "Descriptions.by.Clerk" which contains a code for each occupation.
```{r}
data("almshouse_admissions")
head(almshouse_admissions$Descriptions.by.Clerk)
```
For the purposes of working with this data in R, having only the code isn't very useful. The code book for these occupations is available here: 
```{r}
almshouse.occupations <- read.csv(file="https://raw.githubusercontent.com/regan008/DigitalMethodsData/main/raw/almshouse-occupationalcodes.csv", header=TRUE)
```

A join allows us to join these two dataframes together, matching each row based on the occupational code provided in the `Descriptions.by.Clerk` column. To do that we'll use a function known as a mutating join. A mutating join allows you to combine variables from two tables. It first matches observations by their keys, then copies across variables from one table to the other. In this case we want to join the matching rows from `almshouse.occupations` to `almshouse_admissions`. In an ideal world, the column names in the two data frames would match but since that isn't the case, we'll have to specify what columns `left_join` should use to join the two data frames. 

```{r}
almshouse_admissions <- left_join(almshouse_admissions, almshouse.occupations, by=c("Descriptions.by.Clerk"="code"))

head(almshouse_admissions)
```

(@) Below I've downloaded data about each of the census regions. Join this dataset with `gayguides`. Create a data frame that includes each of the regions and the total number of locations in 1980. How many locations appear in the Mountain region in 1980?
```{r}
regions <- read.csv("https://raw.githubusercontent.com/regan008/DigitalMethodsData/main/raw/censusregions.csv")

regional_gayguides <- left_join(gayguides, regions, by = c("state" = "State.Code"))


```

(@) Explain what you did above. What variable did you join by and why? What results?
State and  State.code. It creates a new dataset that includes everything in the gayguides whilst adding regions and division. 

(@)How much did LGTBQ life grow between 1970 and 1980? Can you create a data frame that computes the growth in the number of locations between 1970 and 1980 for every state? For every region? 
```{r}
# State Data Frame
state_1970 <- regional_gayguides |> 
  filter(Year == 1970) |> 
  select(Year, State) |> 
  group_by(State) |>
  summarize(state_count_1970 = n())
state_1980 <- regional_gayguides |> 
  filter(Year == 1980) |> 
  select(Year, State) |> 
  group_by(State) |>
  summarize(state_count_1980 = n())
state_gg <- left_join(state_1970, state_1980, by = c("State" = "State"))
state_gg <- state_gg |> mutate(Growth = (state_count_1980 - state_count_1970))

region_1980 <- regional_gayguides |>
  filter(Year == 1980) |> 
  select(Year, Region) |> 
  group_by(Region) |>
  summarize(region_count_1980 = n())

#Regional data frame
region_1970 <- regional_gayguides |>
  filter(Year == 1970) |> 
  select(Year, Region) |> 
  group_by(Region) |>
  summarize(region_count_1970 = n())
regional_gg <- left_join(region_1970, region_1980, by = c("Region" = "Region"))
regional_gg <- regional_gg |> mutate(Regional_Growth = (region_count_1980 - region_count_1970))

```


### `pivot_longer()` and `pivot_wider()`: Converting Wide and Long Data

It's possible that you won't create every dataset you use in R. Sometimes that means the dataset is in a format that isn't useful for the questions you want to ask. The dataset below is what is referred to as a "wide" data frame. That is in comparison to a "long" data frame (which would be considered tidy data).
```{r}
library(tidyr)
sc.parks <- read.csv("https://raw.githubusercontent.com/regan008/DigitalMethodsData/main/raw/RecreationData-Wide.csv")
head(sc.parks)
```
This dataset contains all of the localities in South Carolina along with information about the types of recreational workers in that city (paid vs unpaid, male vs female). However, the problem with this dataset is that every year is a column heading making it difficult to work with. On the surface this seems like a useful format, partially because it reads left to right which is how we're accustomed to reading documents. Its easy to compare, for example, the number of female paid recreation workers between 1930 and 1945. But for computational purposes this format is less than ideal for many types of visualizations and operations. R provides functions for dealing with this. `pivot_longer()` "lengthens" your data by increasing the number of rows and decreasing the number of columns. 
```{r}
sc.parks <- sc.parks %>%
  pivot_longer(!city:type_of_worker, names_to = "year", values_to = "count")
```

(@) What did this code do? 
> Takes all the years and puts them as the cells for the new column "year." Additionally, count is the amount below and assigned to the original "xyear" columns. 

(@) Here's another wide data frame. Can you turn this from a wide to a narrow data frame? 
```{r}
rec.spaces <- read.csv("https://raw.githubusercontent.com/regan008/DigitalMethodsData/main/raw/PlayAreabyType.csv")

rec_spaces <- rec.spaces |> pivot_longer(!type, names_to = "year", values_to = "count")
```

The opposite of `pivot_longer()` is `pivot_wider()`. It "widens" data by increasing the number of columns and decreasing the number of rows. We can revert `sc.parks` back to a wide dataset using this function.
```{r}
sc.parks %>%
  pivot_wider(names_from = year, values_from = count)
```

(@) Widen the `sc.parks` dataset so that the column names are drawn from the type of recreation worker.
```{r}
sc <- sc.parks |> pivot_wider(names_from = type_of_worker, values_from = count)
```

(@) Turn `rec.spaces` into a wide dataframe. 
```{r}
rec.spaces |> pivot_wider(names_from = type, values_from = count)

```

## Working with Factors using `forcats`

You were introduced to factors at the end of Worksheet 4. Factors are R's way of handling categorical data - things like occupation, location type, region, etc. The `forcats` package (part of tidyverse) provides tools for working with factors.

### Why Factors Matter for Historians

Historical data is full of categories:
- Occupations (farmer, merchant, laborer...)
- Location types (bar, restaurant, bookstore...)
- Geographic regions (Northeast, South, West...)

Factors let you:
- Control the order categories appear in tables and plots
- Combine rare categories
- Rename categories for clarity
- Ensure consistency

### Creating and Examining Factors

Let's work with the Boston Women Voters data and look at occupations:
```{r}
data("BostonWomenVoters")

# Look at the occupations
head(BostonWomenVoters$Occupation)

# Convert to factor and see the levels
BostonWomenVoters$Occupation <- as.factor(BostonWomenVoters$Occupation)
levels(BostonWomenVoters$Occupation)
```

(@) How many different occupations are there?

> 1137 occupations

(@) Look at the levels. Do you notice any that could be combined or standardized? List a few examples.

> Yes. The different ways Waitress has been spelt in the dataset can be fixed. Also, tel. operator and other versions of telephone operator can be standardizered. Teachers can also be standardized based on their genre. 

### `fct_count()` - Counting Factor Levels

Before reordering or combining categories, it's useful to see how many of each you have:
```{r}
BostonWomenVoters %>%
  pull(Occupation) %>%
  fct_count() %>%
  arrange(desc(n))
```

(@) What are the three most common occupations?

> Housewife, At home, and Housekeeper. 

### `fct_lump()` - Combining Small Categories

When you have many categories and some are rare, `fct_lump()` combines the smallest categories into "Other":
```{r}
BostonWomenVoters %>%
  mutate(occupation_lumped = fct_lump(Occupation, n = 5)) %>%  # Keep top 5, lump the rest
  count(occupation_lumped) %>%
  arrange(desc(n))
```

This kept the 5 most common occupations and combined everything else into "Other".

(@) Use `fct_lump()` to keep the top 10 occupations. How many voters are in the "Other" category?
```{r}
BostonWomenVoters |>
  mutate(occupation_lumped = fct_lump(Occupation, n = 10))|>
  count(occupation_lumped) |>
  arrange(desc(n))
##4759
```

### `fct_reorder()` - Reordering by Another Variable

By default, factors are ordered alphabetically. But often you want them ordered by frequency or some other value. This is especially useful for plotting which we will get to in the next worksheet.
```{r}
# Count by occupation
occupation_counts <- BostonWomenVoters %>%
  count(Occupation) %>%
  filter(n > 30)  # Only occupations with more than 30 people

```
### `fct_reorder()` with Other Variables

You can also reorder by variables other than count. For example, order occupations by average age:
```{r}
avg_age_Boston <- BostonWomenVoters %>%
  group_by(Occupation) %>%
  summarize(avg_age = mean(Age, na.rm = TRUE),
            count = n()) %>%
  filter(count > 20) %>%  # Only occupations with at least 20 people
  mutate(Occupation = fct_reorder(Occupation, avg_age)) |> arrange(avg_age)

head(avg_age_Boston)
tail(avg_age_Boston)

```

(@) Which occupations have the youngest average age? Which have the oldest? What might this tell us historically?

> Students, telephone operators, and school teachers have the youngest average age. The occupations for the oldest women are no employment, at home, matron, and housekeeper. Students are normally younger on average, operators may have been a lower entry job, and school teachers is an interesting job available to these younger women. As for older, most of these women may be at an age where they are unable to work a job, retired, or only able to do what their old age allows them to do.

### `fct_recode()` - Renaming Factor Levels

Sometimes you need to rename categories for clarity or to fix inconsistencies:
```{r}
BWV.occupations <- BostonWomenVoters %>%
    mutate(
      Occupation_clean = fct_recode(Occupation,
        "at home" = "At Home",
        "at Home" = "At Home",
        "Book-keeper" = "Bookkeeper"
      )
    ) %>%
    count(Occupation_clean) %>%
    arrange(desc(n))
```

(@) Look at the occupation levels. Find at least 3 more occupations that could be recoded to be more standardized or descriptive. Create a new column with your recoded occupations.
```{r}
 new_BWV_occupations <- BostonWomenVoters %>%
    mutate(
      Occupation_cleaned = fct_recode(Occupation,
        "School teacher" = "School Teacher",
        "waitress" = "Waitress",
        "Telephone operator" = "Telephone Operator"
      )
    )
```

### `fct_collapse()` - Combining Multiple Levels into Categories

`fct_collapse()` is like `fct_recode()` but designed for grouping many categories into broader themes:
```{r}
test <- BostonWomenVoters %>%
  mutate(
    occupation_category = fct_collapse(Occupation,
      "Domestic Work" = c("Housewife", "At Home", "Housekeeper"),
      "Education" = c("Teacher", "School Teacher"),
      "Skilled Trade" = c("Dressmaker", "Seamstress", "Milliner"),
      "Clerical" = c("Clerk", "Stenographer", "Bookkeeper"),
      "Sales" = c("Saleswoman", "Sales")
    )
  ) %>%
  count(occupation_category) %>%
  arrange(desc(n))
```
Note that this dataset is still really long. We'd have to add all the categories to clean this up fully. But if you look, you'll notice that the new categories are included near the top.

(@) Create your own categorization scheme for occupations. Combine related occupations into 5-7 broader categories that make historical sense. Consider categories like professional, domestic, skilled labor, etc. Explain your reasoning.
```{r}
categorical_BWV <- BostonWomenVoters |>
  mutate(
    occupation_category = fct_collapse(Occupation,
      "Domestic" = c("Housewife", "At Home", "Housekeeper", "Domestic", "Maid", "domestic" ),
      "Education" = c("Teacher", "School Teacher", "Student", "Medical Student", "student", "teacher"),
      "Unemployed" = c("Unemployed", "unemployed", "none", "None"), 
      "Arts" = c("musician", "Music Teacher", "Music teacher", "Actress", "actress", "Artist", "artist", "Musician"),
      "Medical" = c("Nurse", "nurse", "Graduate Nurse", "Trained nurse", "Dietitian", "dental hyginnist", "Physician", "physician" ),
      "Fashion" = c("Dressmaker", "Seamstress", "Milliner", "Hairdresser", "dressmaker", "Shoe worker", "Tailoress", "dress-maker")
    )
  ) |>
  count(occupation_category) |>
  arrange(desc(n))
```

> Explain your reasoning. 
## The category "Domestic" follows jobs revolving around domestic labor. The category "Education" follows teachers and students who are both either educating or being educated at the time. "Unemployed" is a category for women who have no job. The "Arts" category is for women whose jobs are primarily within the arts. The "Medical" category are the jobs within the medical sphere. Lastly, the "Fashion" category is for women who hold jobs relating in some sense to fashion. 

### `fct_relevel()` - Manual Reordering

Sometimes you want a specific order that isn't alphabetical or by frequency. For example, you might want to order occupations by social/economic status to show income levels on a graph:
```{r}
BostonWomenVoters %>%
  mutate(
    occupation_category = fct_collapse(Occupation,
      "Professional" = c("Teacher", "Nurse"),
      "Clerical" = c("Clerk", "Stenographer"),
      "Domestic" = c("Housewife", "At Home")
    ),
    # Order by typical class hierarchy
    occupation_category = fct_relevel(occupation_category, 
                                     "Professional", "Clerical", "Domestic")
  ) %>%
  pull(occupation_category) %>%
  levels()
```

Now when you use this factor in a plot or table, it will appear in this conceptual order rather than alphabetically.

(@) Create occupation categories and relevel them in an order that makes sense historically. Explain why you chose this order.
```{r}
BostonWomenVoters |>
  mutate(
    occupation_category = fct_collapse(Occupation,
      "Domestic" = c("Housewife", "At Home", "Housekeeper", "Domestic", "Maid", "domestic" ),
      "Education" = c("Teacher", "School Teacher",  "teacher"),
      "Unemployed" = c("Unemployed", "unemployed", "none", "None"), 
      "Arts" = c("musician", "Music Teacher", "Music teacher", "Actress", "actress", "Artist", "artist", "Musician", "Writer"),
      "Medical" = c("Nurse", "nurse", "Graduate Nurse", "Trained nurse", "Dietitian", "dental hyginnist", "Physician", "physician" ),
      "Fashion" = c("Dressmaker", "Seamstress", "Milliner", "Hairdresser", "dressmaker", "Shoe worker", "Tailoress", "dress-maker")
    ),
   occupation_category = fct_relevel(occupation_category, 
                                     "Medical", 
                                     "Education",
                                     "Fashion", 
                                     "Arts", 
                                     "Domestic", 
                                     "Unemployed")
  )|> 
  pull(occupation_category) |> 
  levels()
```

> Explanation:
I wanted to make the occupations in terms of a combined interest in the level of education and probable job security each job required and offered. I wanted to include unemployed at the bottom as the basis for what the worst can be. 

## Working with Dates using `lubridate`

Dates in historical data are notoriously messy. They come in different formats, sometimes incomplete, and often require calculation (age, duration, etc.). The `lubridate` package makes working with dates much easier.

### Parsing Dates from Strings

`lubridate` has functions named after date order:
- `mdy()` - month-day-year (American format): "12/25/2020"
- `dmy()` - day-month-year (European format): "25/12/2020"  
- `ymd()` - year-month-day (ISO format): "2020-12-25"
```{r}
# Different date formats
library(lubridate)
mdy("12/25/2020")
dmy("25-12-2020")
ymd("2020-12-25")

# Also work with times
mdy_hms("12/25/2020 14:30:00")
```

These all create "Date" objects that R knows how to work with.

(@) Create a vector of dates in different formats and parse them using the appropriate lubridate functions.
```{r}
date <- c("12/12/2024", "01-12-2020", "2025-01-12", "January 21, 1993")
mdy(date)
dmy(date)
ymd(date)
```

### Creating Date Columns

Let's create some sample historical data:
```{r}
events <- data.frame(
  event = c("Election", "Treaty Signed", "Battle"),
  date_string = c("November 5, 1920", "June 28, 1919", "July 1, 1863")
)

events %>%
  mutate(date = mdy(date_string))
```

(@) The `gayguides` data has a Year column but no full dates. Create a date column assuming all entries are from January 1st of that year. (Hint: use `paste()` to create a string like "1970-01-01", then parse it)
```{r}
date_gayguides <- gayguides |> mutate(
  date = paste(Year, "01", "01", sep="-")
)
date_gayguides |> mutate(date = ymd(date))
```

### Extracting Components from Dates

Once you have a proper date, you can pull out pieces:
```{r}
dates <- ymd(c("1920-11-05", "1919-06-28", "1863-07-01"))

year(dates)
month(dates)
day(dates)
month(dates, label = TRUE)  # Get month names
wday(dates, label = TRUE)   # Get day of week
```

(@) Using the dates you created above for `gayguides`, extract the decade from the date. Compare it to using `floor(Year / 10) * 10` directly on the Year column. Do you get the same result?
```{r}
date_gayguides2 <- date_gayguides |> mutate (decade = floor(year(date)/10) * 10)
d_gayguides <- gayguides |> mutate(decade = floor(Year/10) * 10)
## Yes. 
```

### Date Arithmetic

You can do math with dates:
```{r}
start_date <- ymd("1920-01-01")
end_date <- ymd("1925-12-31")

# Time between dates
end_date - start_date

# Add/subtract time
start_date + days(30)    # 30 days later
start_date + years(1)    # 1 year later
start_date + months(6)   # 6 months later
```

(@) The Boston Women Voters data was collected in 1920. Each person has an age. Calculate each person's approximate birth year using `1920 - Age`. Then create a proper birth date assuming they were born on July 1st of that year.
```{r}
Age_BVW <- BostonWomenVoters |> mutate(
  birthyear = 1920 - Age) |> mutate(
  birthdate = paste(birthyear, "06", "01", sep="-")
)
```

### Working with Intervals
```{r}
# Create an interval
prohibition <- interval(ymd("1920-01-17"), ymd("1933-12-05"))

# How long was prohibition?
prohibition / years(1)   # In years
prohibition / months(1)  # In months

# Was a date during prohibition?
ymd("1925-06-15") %within% prohibition  # TRUE
ymd("1935-06-15") %within% prohibition  # FALSE
```

(@) Create intervals for different decades (1960s, 1970s, 1980s). For each entry in `gayguides`, determine which decade interval it falls into.
```{r}
sixties <- interval(ymd("1960-01-01"), ymd("1969-01-01"))
seventies <- interval(ymd("1970-01-01"), ymd("1979-01-01"))
eighties <- interval(ymd("1980-01-01"), ymd("1989-01-01"))
decade_gayguides <- function(date) {
  data(date_gayguides)
  for (i in 1:nrow(date_gayguides)) {
    if (date_gayguides$date[i] == x && ymd("x") %within% sixties) } {
    print("sixties")
    } else if (date_gayguides$date[i] == x && ymd("x") %within% seventies) {print("seventies")}{ 
      else if(date_gayguides$date[i] == x && ymd("x") %within% eighties) 
    } 
   else 
}
????
```

### Handling Historical Dates

Historical dates can be tricky when month and day might unknow or not exact. Computers demand specificity and a level of certainty and often times historical sources do not support this. 

For dates where day/month are unknown, you can use the first of the year/month:
```{r}
# If you only know the year
year_only <- ymd(paste0("1920", "-01-01"))

# If you know year and month
year_month <- ymd(paste0("1920-06", "-01"))
```

(@) Some historical records might only have a year. Create a function that takes a year and returns a date object for January 1st of that year. Test it with several years.
```{r}
jan_year <- function(x) {ymd(paste(x, "-01-01"))}
jan_year(1960)
jan_year(2000)
```

(@) Reflect: What challenges might you encounter when working with dates in historical sources? How might `lubridate` help address some of these challenges?

> Some issues you might face is that dates could be in different formats, or leave out one or more of day, month, and year. Luridate helps address these issues by allowing in easy way to fix any sources missing data, and by easily fixing dates to one form by using a function. 

## Putting it all together
Each of the functions covered in this worksheet and in last week's worksheet are valuable tools for manipulating datasets. But they are more powerful when combined. When using them to pair down a dataset, we are asking and answering a question. For example in this code from earlier in our worksheet:
```{r}
gayguides %>% 
    select(title, Year) %>%
    group_by(Year) %>%
    summarize(count = n())
```
The implicit question was, "How many locations appear in each year?". The `judges` dataset in provided in the DigitalMethodsData package is a large, messy, wide dataframe that contains a lot of information. Look at this dataframe and then compose a question to ask of the data.

(@) First, tell me, what is the question you are asking? 
```{r}
data("judges")
What is the birth year that holds the most female nominations? 

```

(@) Now write some code to address that question. Comment the code with notes that explain your thinking as you go. Use functions like select(), filter(), etc to pair down your dataset and reshape it to address your question. 
```{r}

judge_year <- judges |> 
    ##Filters gender to only find females who were chosen to be nominated. 
    filter(Gender == "Female")|> 
    ## Selects Gender and Birth Year to find the data I need. 
    select(Gender, Birth.Year) |>
    ## Group by birth year to be able to set up the count. 
    group_by(Birth.Year) |>
    ## Finds every individual born in a given year who was nominated. 
    summarize(count = n()) |>
    ## Arranges the years by the amount to easily find the answer. 
    arrange(desc(count))
## Finds the answer. 
head(judge_year, 1)

```

(@) Now ask a question of the `gayguides` data (or another dataset of your choice). What is the question you are asking? 
```{r}
## What are the top 5 states or provinces of birth women in ward 6 are from?
```

(@) Now write some code to address that question. Comment the code with notes that explain your thinking as you go. Use functions like select(), filter(), etc to pair down your dataset and reshape it to address your question. 
```{r}
birth_state_BVW <- BostonWomenVoters |> 
### Filter to only pull ward 6.
    filter((Ward == "6")) |>
    ## Select the ward and state of birth columns. 
    select(Ward, State.or.Province.of.Birth) |>
    # Group by state of birth. 
    group_by(State.or.Province.of.Birth) |>
    ## Count entires. 
    summarize(count = n())|>
    ## Arrange them from highest to lowest. 
    arrange(desc(count))
## Prints the answer. 
head(birth_state_BVW)
```

(@) Write a function that filters the gay guides dataset. It should accept 2 arguments: year and state. When passed to the function the function should return only the title, type, state and year for each entry. 
```{r}
state_year <- function(year, state) {
  data(gayguides)
  for (i in 1:length(gayguides$state)) {
    if (gayguides$Year[i] == year && gayguides$state[i] == state) {
     print(paste(gayguides$title[i], gayguides$type[i], state, year))
    } else {
      next 
    } 
  }
}
state_year("1982", "NY")
```

(@) Lets try another, using the BostonWomenVoters data, answer this question: "How did the age distribution of women in different occupational categories vary across wards?"

Your analysis should:
- Remove any rows with missing Age or Ward data
- Create broader occupation categories using `fct_collapse()` or `fct_lump()`
- Group by Ward and your occupation categories
- Calculate average age, minimum age, and maximum age for each group
- Arrange the results in a meaningful way
```{r}
aa <- BostonWomenVoters |> drop_na(Age, Ward) |> mutate(
    occupation_category = fct_collapse(Occupation,
      "Domestic" = c("Housewife", "At Home", "Housekeeper", "Domestic", "Maid", "domestic", "At home", "Housewife", "Hpusework", "H.W.", "housewife" ),
      "Education" = c("Teacher", "School Teacher", "Student", "Medical Student", "student", "teacher"),
      "Unemployed" = c("Unemployed", "unemployed", "none", "None"), 
      "Arts" = c("musician", "Music Teacher", "Music teacher", "Actress", "actress", "Artist", "artist", "Musician"),
      "Medical" = c("Nurse", "nurse", "Graduate Nurse", "Trained nurse", "Dietitian", "dental hyginnist", "Physician", "physician" ),
      "Fashion" = c("Dressmaker", "Seamstress", "Milliner", "Hairdresser", "dressmaker", "Shoe worker", "Tailoress", "dress-maker")
    )
  ) |> group_by(Ward, occupation_category) |>
  summarize(
    count = n(),
    avg_age = mean(Age, na.rm = TRUE),
    min_age = min(Age, na.rm = TRUE),
    max_age = max(Age, na.rm = TRUE)
  ) |> arrange(desc(count))
```

(@) What did you discover? Write 2-3 sentences about what this tells us about Boston women voters.

> Domestic occupations seemed to be the highest form of occupation in many of the wards, with the top 5 occupations all being domestic in different wards. Ward 8 faced a major unemployment issue, as it was their second biggest occupation  category. Lastly, many of the more professional or skilled jobs see little jobs compared to less skilled jobs, perhaps suggesting the difficulty for women to break into these types of jobs at the time. 

